### 2012-04-09 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: Corrections.

Path expressions without predicates which use only child axis are the simpliest example of such queries. For example, path expression \texttt{/bib/book/author} indicates that element \texttt{bib} is the root element and it contains one or more elements \texttt{book} and these contain one or more elements \texttt{author}.

@@

Statements of this kind are necessary for the XML schema inference, however their obtaining from queries is not significant due to the following reason. These statements could be determined directly from the XML documents and it could be easily done in a more convenient way. In addition, queries may not cover the whole relevant content of the documents. Let \texttt{E} be an element occurring in an XML document and let \texttt{Q} be an XQuery query. Consider the following cases:

\begin{enumerate}[1)]
\item \texttt{E} is directly mentioned in \texttt{Q}. For example elements \texttt{bib}, \texttt{book} and \texttt{author} are mentioned in path expression \texttt{/bib/book/author}.
\item \texttt{E} is not mentioned in \texttt{Q} but it does occur in a result of the evaluation of \texttt{Q}. For example path expression \texttt{/bib/book/author} returns elements \texttt{author} along with their subelements \texttt{name}, \texttt{birthdate} and \texttt{nationality}.
\item \texttt{E} is neither mentioned in \texttt{Q}, nor it occurs in the result, but it is processed by the evaluation of \texttt{Q}. For example the evaluation of path expression \texttt{//author} processes amongst others elements \texttt{bib} and \texttt{book}, in a search for elements \texttt{author}.
\item \texttt{E} occurs in a part of the XML document not related to \texttt{Q} at all.
\end{enumerate}

How we can obtain any information about element \texttt{E} from query \texttt{Q}? In case 1), \texttt{Q} can be directly exploited in the inference process. Case 2) requires the result of \texttt{Q} and additional processing of the result. Case 3) requires a step-by-step evaluation of \texttt{Q}, hence, the XML document is also required. And in case 4), it is not possible to obtain any information.

@@

Basic representant is a common \emph{FLWOR} query.

>>>>

Path expressions without predicates which use only child axis are the simplest example of such queries. For example, path expression \texttt{/bib/book/author} indicates that element \texttt{bib} is the root element and it contains one or more elements \texttt{book} and these contain one or more elements \texttt{author}.

@@

Statements of this kind are necessary for the XML schema inference, however their obtaining from queries is not significant due to the following reason. These statements could be determined directly from the XML documents and it could be easily done in a more convenient way. In addition, queries may not cover the whole relevant content of the documents. Let $E$ be an element occurring in an XML document and let $Q$ be an XQuery query. Consider the following cases:

\begin{enumerate}[1)]
\item $E$ is directly mentioned in $Q$. For example elements \texttt{bib}, \texttt{book} and \texttt{author} are mentioned in path expression \texttt{/bib/book/author}.
\item $E$ is not mentioned in $Q$ but it does occur in a result of the evaluation of $Q$. For example path expression \texttt{/bib/book/author} returns elements \texttt{author} along with their subelements \texttt{name}, \texttt{birthdate} and \texttt{nationality}.
\item $E$ is neither mentioned in $Q$, nor it occurs in the result, but it is processed by the evaluation of $Q$. For example the evaluation of path expression \texttt{//author} processes amongst others elements \texttt{bib} and \texttt{book}, in a search for elements \texttt{author}.
\item $E$ occurs in a part of the XML document not related to $Q$ at all.
\end{enumerate}

How we can obtain any information about element $E$ from query $Q$? In case 1), $Q$ can be directly exploited in the inference process. Case 2) requires the result of $Q$ and additional processing of the result. Case 3) requires a step-by-step evaluation of $Q$, hence, the XML document is also required. And in case 4), it is not possible to obtain any information.

@@

Basic representant is a common FLWOR query.

### 2012-04-09 doc/theses/mikula/diplomka.tex
Master thesis: Corrections and final(?) additions.
Master thesis: Corrections.

\hbox{\hbox to 0.5\hsize{%
In ........ date ............
\hss}\hbox to 0.5\hsize{%
Signature
\hss}}

@@

Abstrakt: V poslednej dobe bolo vyvinutých množstvo metód zaoberajúcich sa automatickým odvodzovaním XML schémy. Väčšina z nich ale používa XML dokumenty ako jediný vstup. My sa v tejto práci zameriavame na rozšírenie odvo\-dzo\-vania zahrnutím XML operácií, konkrétne XQuery dotazov. Diskutujeme ako je možné využiť XQuery dotazy na vylepšenie procesu odvodzovania a navrhujeme algoritmus založený na vybraných vylepšeniach, ktorý rozširuje existujúcu metódu hľadania kľúčov a ktorý môže byť začlenený do tzv. \emph{metód odvodzujúcich gramatiku}. Implementovaním algoritmu sme vytvorili prvé kompletné riešenie odvodzovania XML schémy, ktoré využíva XML dokumenty spolu s XML operáciami.
% abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce

@@

Abstract: Recently, plenty of methods dealing with automatic inference of XML schema have been developed, however, most of them utilize XML documents as their only input. In this thesis we focus on extending inference by incorporating XML operations, in particular XQuery queries. We discuss how can XQuery queries help in improving the inference process and we propose an algorithm based on chosen improvements, extending an existing method of a key discovery, that can be integrated to so-called \emph{grammar-inferring} methods. By implementing it, we created the first solution of XML schema inference using XML documents along with XML operations.
% abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad
% zadání diplomové práce

@@

%%% Tabulky v diplomové práci, existují-li.
%\chapwithtoc{List of Tables}
\listoftables

@@

%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
%%% diagramy apod.). Kaľdá příloha musí být alespoň jednou odkazována z vlastního
%%% textu práce. Přílohy se číslují.
%\chapwithtoc{Attachments}
\appendix
\include{appendix_a}
\include{appendix_b}

>>>>

\hbox{\hbox to 0.5\hsize{%
In ........ date ............
\hss}\hbox to 0.5\hsize{%
%Signature
\hss}}

@@

Abstrakt: V poslednej dobe bolo vyvinutých množstvo metód zaoberajúcich sa automatickým odvodzovaním XML schémy. Väčšina z nich ale používa XML dokumenty ako jediný vstup. My sa v tejto práci zameriavame na rozšírenie odvo\-dzo\-vania zahrnutím XML operácií, konkrétne XQuery dotazov. Diskutujeme ako je možné využiť XQuery dotazy na vylepšenie procesu odvodzovania a navrhujeme algoritmus založený na vybraných vylepšeniach, ktorý rozširuje existujúcu metódu hľadania kľúčov a ktorý môže byť začlenený do metód odvodzujúcich tzv. \emph{počiatočnú gramatiku}. Implementovaním algoritmu sme vytvorili prvé kompletné riešenie odvodzovania XML schémy, ktoré využíva XML dokumenty spolu s XML operáciami.
% abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce

@@

Abstract: Recently, plenty of methods dealing with automatic inference of XML schema have been developed, however, most of them utilize XML documents as their only input. In this thesis we focus on extending inference by incorporating XML operations, in particular XQuery queries. We discuss how can XQuery queries help in improving the inference process and we propose an algorithm based on chosen improvements, extending an existing method of a key discovery, that can be integrated to methods inferring so-called \emph{initial grammar}. By implementing it, we created the first solution of XML schema inference using XML documents along with XML operations.
% abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad
% zadání diplomové práce

@@

@@

%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
%%% diagramy apod.). Kaľdá příloha musí být alespoň jednou odkazována z vlastního
%%% textu práce. Přílohy se číslují.
%\chapwithtoc{Attachments}
\appendix
\include{appendix_a}
\include{appendix_b}
\include{appendix_c}

%%% Tabulky v diplomové práci, existují-li.
%\chapwithtoc{List of Tables}
\listoftables
\listoffigures

### 2012-04-09 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: Corrections.

\begin{figure}
\begin{alltt}
<xs:element name="book">
  <xs:complexType>
    <xs:sequence>
      <xs:element name="title" type="xs:string"/>
      <xs:element name="authors>
        <xs:complexType>
          <xs:element name="author" type="xs:string"
           maxOccurs="unbound"/>
        </xs:complexType>
      </xs:element>
      <xs:element name="isbn" type="xs:string" minOccurs="0"/>
    </xs:sequence>
    <xs:attribute name="id" type="xs:string" use="required"/>
  </xs:complexContent>
</xs:element>
\end{alltt}
\caption{A simple XSD example}
\label{FIG_a_simple_xsd_example}
\end{figure}

>>>>

\begin{figure}
\begin{alltt}
<xs:element name="book">
  <xs:complexType>
    <xs:sequence>
      <xs:element name="title" type="xs:string"/>
      <xs:element name="authors">
        <xs:complexType>
          <xs:sequence>
            <xs:element name="author" type="xs:string"
             maxOccurs="unbounded"/>
          </xs:sequnce>
        </xs:complexType>
      </xs:element>
      <xs:element name="isbn" type="xs:string" minOccurs="0"/>
    </xs:sequence>
    <xs:attribute name="id" type="xs:string" use="required"/>
  </xs:complexType>
</xs:element>
\end{alltt}
\caption{A simple XSD example}
\label{FIG_a_simple_xsd_example}
\end{figure}
